### 2022-03-11 paper/paper.tex
fix notation error in forward pass
Shorten title
Switch to one sentence per line for better version control

    \definecolor{ul-red}{RGB}{220,29,39}
    \begin{figure}[htb!]
        \centering
        \includegraphics[width=0.1\linewidth]{raster/ul-logo}
    \end{figure}
    \huge{\textsc{University of Ljubljana}}\\
    \LARGE{\textsc{Faculty of {\color{ul-red} Mathematics and Physics}}}\\[1mm]
    \Large{\textsc{Department of Physics}}\\
    \vspace{5mm}
    \huge{Seminar}\\
    \vspace{5mm}
    \rule{\textwidth}{0.2pt}\\[3mm]
    \LARGE{\textbf{End-to-End Classification for Discovery of\\New Processes in High-Energy Physics}}
    % End-to-End Particle Classification for Discovery of New Processes in High-Energy Physics
    \rule{\textwidth}{0.2pt}

@@

    % \large{\textit{Prepared by}: Elijan Jakob Mastnak}\\
    % \large{\textit{Adviser}: prof. dr. Borut Paul Ker\v{s}evan}

@@

\vspace{10mm}
% add myurl links to all sources
\begin{center}
    \textbf{Abstract}\\[2mm]
    \begin{minipage}[t]{0.85\textwidth}
        This work explores the use of convolutional neural networks to classify the products of high-energy particle collisions using low-level, image-based detector data. We restrict ourselves to binary classification, and, for concreteness, focus on the popular example of distinguishing Higgs boson and background events. We begin by defining end-to-end classification, explain how the data used to perform this classification is measured at the Large Hadron Collider, and outline how both fully-connected and convolutional neural networks use this data to perform classification. We conclude by presenting a concrete study involving image-based end-to-end classification, and compare end-to-end workflows to traditional classification methods based on reconstructed kinematic features.
    \end{minipage}

@@

 As implied by Figures \ref{fig:traditional-pipeline} and \ref{fig:cnn-pipeline}, particle classification requires data measured by a particle detector. In the next section, we explore the nature of this detector data, and outline how the data is produced and measured at the Large Hadron Collider.

@@

% Pooling acts only along spatial dimensions, and preserves the channel dimension. For any input and pooling kernel dimensions $ I_{j} $, $ K_{j}^{(\text{p})} $ and any kernel stride $ s_{j} $, the output size of a pooling layer along the axis $ j $ is
% \begin{equation}
%     O_{j}^{(\text{p})} = \floor*{\frac{I_{j} - K_{j}^{(\text{p})}}{s_{j}}} + 1; \label{eq:cnn-O-pool}
% \end{equation}
% this is simply Equation \ref{eq:cnn-O-conv} with $ p_{j} = 0 $, since pooling does not involve zero padding.

@@

%     % \item Define input images with shape $ C_{\text{in}}, I_{\text{w}}, I_{\text{h}} $. Define input layer. This is a convolutional layer.

%     % Define an arbitrary number of filters to use in the input layer. Each filter has dimensions $ (C_{\text{in}}, K_{\text{w}}, I_{\text{h}}) $. Use $ C_{\text{out}} $ such filters. Perform discrete convolutions of an input image with each filter as described in \myhref{ss:discrete-conv}{Section \ref{ss:discrete-conv}} and shown Figure \ref{fig:conv-multi-channel}. The result is an output with shape $ (C_{\text{out}}, O_{\text{w}}, O_{\text{h}}) $, where $ O_{\text{w}} $ and $ O_{\text{h}} $ as described in \myhref{ss:conv-arithmetic}{Section \ref{ss:conv-arithmetic}}.

@@

\end{itemize}
% \textit{Note}: The following two sections contain a lot of information. From personal experience, keeping track of the relationships between weights and biases across network layers may be overwhelming to the unfamiliar reader on a first attempt. Unfortunately, I know of no way around this if we wish to meaningfully understand a neural network beyond mere hand-waving. To make things easier, I have paid utmost care to consistent and semantically meaningful notation, while a more intuitive ``big-picture'' interpretation of a neural network follows in \myhref{ss:fcn-interpretation}{Section \ref{ss:fcn-interpretation}}.

@@

Moving forward through the network, the first hidden layer's activation values $ \a^{(1)} \in \mathbb{R}^{n_{1}} $ are passed forward to each of the $ n_{2} $ neurons in the second hidden layer. The pre-activation values of the second hidden layer's neurons are then
\begin{equation*}
    \z^{(2)} = \big( \W^{(2)} \big)^{\mspace{-4mu}\top} \z^{(1)} + \b^{(2)} \in \mathbb{R}^{n_{2}},
\end{equation*}
while the activation values of the neurons in the second hidden layer are
\begin{equation*}
    \a^{(2)} = f_{\text{a}}(\z^{(2)}) = f_{\text{a}} \Big( \big( \W^{(2)} \big)^{\mspace{-4mu}\top} \z^{(1)} + \b^{(2)}\Big) \in \mathbb{R}^{n_{2}}.
\end{equation*}
More generally, the activation values of the $ n_{l} $ neurons in the $ l $-th hidden layer are
\begin{equation*}
    \a^{(l)} = f_{\text{a}}(\z^{(l)}) = f_{\text{a}} \Big( \big( \W^{(l)} \big)^{\mspace{-4mu}\top} \z^{(l-1)} + \b^{(l)}\Big) \in \mathbb{R}^{n_{l}}.
\end{equation*}

\subsection{Optimization} \label{ss:optimization}

\subsubsection{Loss} \label{sss:loss}
To review from the brief discussion in \myhref{ss:optimization-brief}{Section \ref{ss:optimization-brief}}, for a classifier with $ C $ classes, loss is given by a scalar-valued \textit{loss function} $ L : \mathbb{R}^{C} \to \mathbb{R} $, which takes as input a predicted classification score $ \hat{\y} $ and label vector $ \y $, and returns a scalar value $ L $ encoding the difference between $ \hat{\y} $ and $ \y $. In practice, for computational efficiency during optimization, loss is computed on the predictions from a \textit{mini-batch} of $ M \sim 128 $ feature vectors instead of individual instances. In this case, the loss generalizes to 
    \begin{equation}
        \mathcal{L} = \frac{1}{M} \sum_{m = 1}^{M} L_{m} (\hat{\y}_{m}; \y_{m}), \label{eq:loss-batch}
    \end{equation}
    which is just the average of the losses associated with each instance in the mini-batch.

>>>>

    \definecolor{ul-red}{RGB}{220,29,39}
    \begin{figure}[htb!]
        \centering
        \includegraphics[width=0.1\linewidth]{raster/ul-logo}
    \end{figure}
    \huge{\textsc{University of Ljubljana}}\\
    \LARGE{\textsc{Faculty of {\color{ul-red} Mathematics and Physics}}}\\[1mm]
    \Large{\textsc{Department of Physics}}\\
    \vspace{5mm}
    \huge{Seminar}\\
    \vspace{5mm}
    \rule{\textwidth}{0.2pt}\\[3mm]
    \LARGE{\textbf{End-to-End Particle Classification\\in High-Energy Physics}}
    \rule{\textwidth}{0.2pt}

@@

@@

\vspace{10mm}
\begin{center}
    \textbf{Abstract}\\[2mm]
    \begin{minipage}[t]{0.85\textwidth}
        This work explores the use of convolutional neural networks to classify the products of high-energy particle collisions using low-level, image-based detector data. We restrict ourselves to binary classification, and, for concreteness, focus on the popular example of distinguishing Higgs boson and background events. We begin by defining end-to-end classification, explain how the data used to perform this classification is measured at the Large Hadron Collider, and outline how both fully-connected and convolutional neural networks use this data to perform classification. We conclude by presenting a concrete study involving image-based end-to-end classification, and compare end-to-end workflows to traditional classification methods based on reconstructed kinematic features.
    \end{minipage}

@@

 As implied by Figures \ref{fig:traditional-pipeline} and \ref{fig:cnn-pipeline}, particle classification requires as input the data measured by a particle detector.
 In the next section, we explore the nature of this detector data, and outline how the data is produced and measured at the Large Hadron Collider.

@@

% Pooling acts only along spatial dimensions, and preserves the channel dimension.
% For any input and pooling kernel dimensions $ I_{j} $, $ K_{j}^{(\text{p})} $ and any kernel stride $ s_{j} $, the output size of a pooling layer along the axis $ j $ is
% \begin{equation}
%     O_{j}^{(\text{p})} = \floor*{\frac{I_{j} - K_{j}^{(\text{p})}}{s_{j}}} + 1; \label{eq:cnn-O-pool}
% \end{equation}
% this is simply Equation \ref{eq:cnn-O-conv} with $ p_{j} = 0 $, since pooling does not involve zero padding.

@@

%     % \item Define input images with shape $ C_{\text{in}}, I_{\text{w}}, I_{\text{h}} $.
    % Define input layer.
    % This is a convolutional layer.

%     % Define an arbitrary number of filters to use in the input layer.
    % Each filter has dimensions $ (C_{\text{in}}, K_{\text{w}}, I_{\text{h}}) $.
    % Use $ C_{\text{out}} $ such filters.
    % Perform discrete convolutions of an input image with each filter as described in \myhref{ss:discrete-conv}{Section \ref{ss:discrete-conv}} and shown Figure \ref{fig:conv-multi-channel}.
    % The result is an output with shape $ (C_{\text{out}}, O_{\text{w}}, O_{\text{h}}) $, where $ O_{\text{w}} $ and $ O_{\text{h}} $ as described in \myhref{ss:conv-arithmetic}{Section \ref{ss:conv-arithmetic}}.

@@

\end{itemize}
% \textit{Note}: The following two sections contain a lot of information.
% From personal experience, keeping track of the relationships between weights and biases across network layers may be overwhelming to the unfamiliar reader on a first attempt.
% Unfortunately, I know of no way around this if we wish to meaningfully understand a neural network beyond mere hand-waving.
% To make things easier, I have done my best to provide consistent and semantically meaningful notation; also, a more intuitive ``big-picture'' interpretation of a neural network follows in \myhref{ss:fcn-interpretation}{Section \ref{ss:fcn-interpretation}}.

@@

Moving forward through the network, the first hidden layer's activation values $ \a^{(1)} \in \mathbb{R}^{n_{1}} $ are passed forward to each of the $ n_{2} $ neurons in the second hidden layer.
The pre-activation values of the second hidden layer's neurons are then
\begin{equation*}
    \z^{(2)} = \big( \W^{(2)} \big)^{\mspace{-4mu}\top} \a^{(1)} + \b^{(2)} \in \mathbb{R}^{n_{2}},
\end{equation*}
while the activation values of the neurons in the second hidden layer are
\begin{equation*}
    \a^{(2)} = f_{\text{a}}(\z^{(2)}) = f_{\text{a}} \Big( \big( \W^{(2)} \big)^{\mspace{-4mu}\top} \a^{(1)} + \b^{(2)}\Big) \in \mathbb{R}^{n_{2}}.
\end{equation*}
More generally, the activation values of the $ n_{l} $ neurons in the $ l $-th hidden layer are
\begin{equation*}
    \a^{(l)} = f_{\text{a}}(\z^{(l)}) = f_{\text{a}} \Big( \big( \W^{(l)} \big)^{\mspace{-4mu}\top} \a^{(l-1)} + \b^{(l)}\Big) \in \mathbb{R}^{n_{l}}.
\end{equation*}

\subsection{Optimization} \label{ss:optimization}
\subsubsection{Loss} \label{sss:loss}
To review from the brief discussion in \myhref{ss:optimization-brief}{Section \ref{ss:optimization-brief}}, for a classifier with $ C $ classes, loss is given by a scalar-valued \textit{loss function} $ L : \mathbb{R}^{C} \to \mathbb{R} $, which takes as input a predicted classification score $ \hat{\y} $ and label vector $ \y $, and returns a scalar value $ L $ encoding the difference between $ \hat{\y} $ and $ \y $.
In practice, for computational efficiency during optimization, loss is computed on the predictions from a \textit{mini-batch} of $ M \sim 128 $ feature vectors instead of individual instances.
In this case, the loss generalizes to
\begin{equation}
    \mathcal{L} = \frac{1}{M} \sum_{m = 1}^{M} L_{m} (\hat{\y}_{m}; \y_{m}), \label{eq:loss-batch}
\end{equation}
which is just the average of the losses associated with each instance in the mini-batch.
